\documentclass[../report.tex]{subfiles}

\begin{document}

\chapter*{ACKNOWLEDGMENT}
\addcontentsline{toc}{chapter}{ACKNOWLEDGMENT}

I would like to express my sincere gratitude to all those who have supported and guided me throughout the completion of this graduation thesis.

First and foremost, I would like to thank my supervisor, Ph.D. NGUYEN VAN B, for his invaluable guidance, continuous support, and insightful feedback throughout my research. His expertise in blockchain technology and software engineering has been instrumental in shaping this project.

I am deeply grateful to the faculty members of the Software Engineering Department at the University of Information Technology for their excellent teaching and for providing me with a solid foundation in computer science and software development.

I would also like to thank the Solana development community and the open-source contributors whose documentation, tutorials, and tools have been essential to the successful implementation of this project.

Special thanks to my family and friends for their unwavering encouragement, patience, and support during my academic journey. Their belief in me has been a constant source of motivation.

Finally, I would like to acknowledge all the research papers, academic publications, and technical resources that have contributed to my understanding of blockchain technology, decentralized applications, and software architecture principles.

\vspace{1cm}

\begin{flushright}
\textit{Ho Chi Minh City, December 2024}\\
\textbf{NGUYEN VAN A}
\end{flushright}

\end{document}
