\begin{center}
    \Large{\textbf{ABSTRACT}}
\end{center}
\vspace{1cm}

The global freelance market, valued at over \$455 billion annually, is dominated by centralized platforms such as Upwork and Fiverr that collect fees ranging from 5\% to 20\%, exercise unilateral control over dispute resolution, and expose participants to payment fraud and arbitrary account suspension. These structural deficiencies stem from the inherent requirement of a trusted third party to hold escrow funds and adjudicate conflicts---a requirement that blockchain technology makes possible to eliminate. Existing decentralized alternatives remain limited in scope, typically addressing only payment transparency while neglecting identity verification, reputation integrity, and platform governance.

This thesis presents FreelanceChain, a fully decentralized freelance marketplace built on the Solana blockchain using the Anchor framework. The proposed system replaces the centralized intermediary with a suite of eighteen interconnected smart contract modules that collectively enforce the entire lifecycle of a freelance engagement: job creation, competitive bidding, milestone-based escrow, work submission, dispute arbitration, and multi-dimensional reputation recording. To address the identity problem inherent in pseudonymous blockchain interactions, FreelanceChain integrates browser-side biometric verification using the face-api.js library, producing a privacy-preserving on-chain Know Your Customer (KYC) record via a dedicated Anchor program. Reputation is encoded as non-transferable Soulbound Tokens (SBTs) that create an immutable, cryptographically verifiable career history. An AI risk assessment module powered by the Groq API and the LLaMA 3.3 70B open-source large language model provides clients with automated multi-dimensional CV analysis, four-score composite risk scoring, and a conversational follow-up advisor. Platform governance is managed through a Decentralized Autonomous Organization (DAO) module with reputation-weighted voting, while a zero-knowledge proof module enables selective credential disclosure without revealing private data.

Deployed on the Solana Devnet with Program ID \texttt{FStwCj8zLzNkz7gTavooDcqF4FSyDcF9o9SCh96yyP7i}, the system achieves sub-second transaction confirmation at a cost below \$0.001 per transaction, demonstrating that blockchain-based freelance intermediation is technically viable, economically practical, and superior to centralized alternatives on dimensions of transparency, security, and participant sovereignty.

\begin{flushright}
\begin{tabular}{@{}c@{}}
Student\\
\textit{(Signature and full name)}\\[1cm]
Nguyen Van Thang
\end{tabular}
\end{flushright}
