\begin{center}
{\Large \textbf{ABSTRACT}}
\end{center}
\vspace{0.5cm}

The online freelance economy is currently dominated by centralized intermediaries that charge service fees in the 5--20\% range, retain unilateral authority over disputes, and expose participants to payment fraud and arbitrary account suspension. Existing decentralized alternatives address only narrow facets of the problem and rarely combine robust identity verification, reputation portability, and community-driven governance within a single system.

This thesis presents \textbf{FreelanceChain}, a proof-of-concept decentralized freelance marketplace implemented on the Solana blockchain. The prototype decentralizes the trust-critical operations of a freelance engagement through a suite of Anchor smart contracts that manage the job lifecycle, dual-track escrow (lump-sum and milestone), dispute handling, and a multi-dimensional reputation system encoded as non-transferable Soulbound Tokens (SBTs). The system also provides an initial implementation of a browser-side biometric KYC pipeline that keeps face descriptors off custodial servers, and an advisory AI subsystem that assists clients in interpreting freelancer profiles. An on-chain DAO module with reputation-weighted voting is included as a prototype of decentralized parameter governance, although the reward-distribution sub-module remains incomplete.

The prototype has been deployed on the Solana Devnet only; no mainnet deployment, large-scale user study, or third-party security audit has been performed. Devnet measurements show sub-second median confirmation and per-transaction costs below USD\,0.001 for the instructions exercised in this work. These results demonstrate the technical feasibility of the proposed architecture on Solana Devnet; broader claims about economic viability and superiority over centralized platforms remain subjects for future work.

\vspace{0.2cm}
\noindent\textbf{Keywords:} decentralized marketplace prototype, Solana, Anchor, smart-contract escrow, Soulbound Token, browser-side biometric verification, DAO governance, large language model.
