This appendix reproduces the questionnaire administered during the requirements-elicitation survey of Section~\ref{subsection:2.1.2}. The survey was distributed through Google Forms between September and November 2025. Respondents were anonymous, participation was voluntary, and only aggregate statistics are reported in Chapter~2.

\section*{Section 1 -- Respondent profile}
\begin{enumerate}[noitemsep,label=Q\arabic*.]
    \item Which role best describes you? (Freelancer / Client / Both)
    \item How many years of experience do you have on online freelance platforms? (0--1 / 1--3 / 3--5 / 5+)
    \item Which platforms have you used most often in the last 12 months? (multi-select: Upwork, Fiverr, Toptal, Freelancer.com, Other)
    \item Approximately how many engagements have you completed in the last 12 months? (1--3 / 4--10 / 11--30 / 30+)
\end{enumerate}

\section*{Section 2 -- Pain points (5-point Likert: Strongly disagree -- Strongly agree)}
\begin{enumerate}[noitemsep,label=Q\arabic*., start=5]
    \item Platform service fees significantly reduce my net income.
    \item Dispute-resolution outcomes on the platforms I use feel fair and impartial. (reverse-coded)
    \item I can easily transfer my reputation between platforms.
    \item I worry that the platform may suspend my account and freeze held balances.
    \item I am able to verify that a counterparty's identity is genuine.
    \item I receive payment reliably and on time.
\end{enumerate}

\section*{Section 3 -- Decentralization familiarity}
\begin{enumerate}[noitemsep,label=Q\arabic*., start=11]
    \item Have you used a Web3 wallet (Phantom, MetaMask, etc.) before? (Yes / No)
    \item Would you be willing to try a decentralized alternative if the fee were below 2\%? (Yes / Maybe / No)
\end{enumerate}

\section*{Section 4 -- Open-ended}
\begin{enumerate}[noitemsep,label=Q\arabic*., start=13]
    \item Briefly describe the worst dispute you have experienced on a freelance platform.
    \item What single feature would most improve your experience?
    \item Which platform behaviours, if any, make you uncomfortable from a privacy perspective?
    \item Other comments you wish to add.
\end{enumerate}

\noindent The aggregated findings, sample size, sampling method, time window, and explicit limitations of this survey are reported in Section~\ref{subsection:2.1.2}.
